---
title: "Scales"
author: "Gina Reynolds, July 2019"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", "ninjutsu"]
    nature:
      highlightLines: yes
      ratio: '16:9'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
knitr::opts_chunk$set(fig.height = 6, out.width = "100%", comment = " ", cache = T, dpi = 300)
source("https://raw.githubusercontent.com/EvaMaeRey/little_flipbooks_library/master/xaringan_reveal_parentheses_balanced.R")
library(gapminder)
library(tidyverse)
```


```{r xaringan-themer, include = FALSE}
library(xaringanthemer)


# xaringanthemer::write_xaringan_theme(text_font_size = ".75cm",
#                                      text_color = "darkblue")

xaringanthemer::mono_accent(
  base_color = "#43418A",
  header_font_google = google_font("Montserrat"),
  text_font_google   = google_font("Montserrat", "200", "200i"),
  code_font_google   = google_font("Droid Mono"),
  text_font_size = ".7cm",
  code_font_size = ".7cm")
  

# write_xaringan_theme(
#   extra_css = list(
#     ".box-right" = list(
#       "height" = "300px",
#       "width" = "300px",
#       "position" = "absolute",
#       "top" = "33%",
#       "left" = "55%"
#     )
#   )
# )

```




---

name: scales

## Modifiers: Asking for modification with scales\_\*\_\* 

> Please, sir, can I have some more?

With the aes() statement.

But we can then go back and ask for more with scales\_\* statements.  These statements provide a quick way to update our preferences for all the mapped aesthetics.   

Start with an scales_x_* and look at the variety of scale options associated with the x position aesthetic.  


---

## Scales


First let's have look at a variety of scaling options that come with the x position aesthetic:  

- scale_x_log10()
- scale_x_reverse()
- scale_x_sqrt()


---


```{r x_scales, eval = F, echo = F}
gapminder %>% 
  filter(year == 2002) %>% 
  ggplot() +
  aes(x = gdpPercap) +
  aes(y = lifeExp) +
  geom_point() +  # this much is familiar #REVEAL
  scale_x_log10() + # overrides default  #REVEAL
  scale_x_reverse() + # overrides log10  #REVEAL
  scale_x_sqrt() + # overrides reverse  #REVEAL
  scale_x_continuous(
                    position = "top",
                    
                     expand = c(0,0),
                     name = "hi", 
                     minor_breaks = 35000,
                     breaks = 2:4*10000,
                     labels = 1:3,
                     ) 
```

`r apply_reveal("x_scales", user_reveal_defined = T)`


---

### More aesthetic scale modifications

Analogously, we can use scale\_\*\_\* statements to modify other aesthetic scales.  Here are some examples

<!-- I think people find it easy to think about x and y position and adjusting those scales, and related adjustment like labs() and scale_*, but we teachers need to help extend these logics to the other aesthetics.  -->


- scale_x_log10() 
- scale_y_reverse(limits = c(100, 0)) 
- scale_color_viridis_d() 
- scale_shape_discrete(solid = F) 
- scale_size_continuous(breaks = c(100, 1000, 10000, 100000)) 
- scale_alpha(range = c(.5, 1)) # full transparency is 0


Let's tour several of those to get a taste.  

---

```{r scales, eval = F, echo = F}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  scale_x_continuous(name = "Engine displacement (L)") +
  scale_y_continuous(name = "Highway MPG") + 
  labs(x = NULL, y = NULL) +
  scale_x_continuous(limits = c(2, 6)) +
  scale_x_continuous(limits = c(0, 10)) + 
  scale_x_continuous(breaks = c(2, 4, 6)) +
  scale_x_continuous(
  breaks = c(2, 4, 6),
  label = c("two", "four", "six")
```  
  
`r apply_reveal("scales")`

---
  
```{r scales2, eval = F, echo = F}
set.seed(2019)
data.frame(
  x = rnorm(10) * 100000,
  y = seq(0, 1, length.out = 10)
) %>% 
  ggplot() +
  aes(x, y) + 
  geom_point() +
  scale_y_continuous(labels = scales::percent) + # scales package
  scale_y_continuous(labels = scales::dollar) + 
  scale_x_continuous(labels = scales::comma) + 
  scale_y_continuous(trans = scales::reciprocal_trans())
```
`r apply_reveal("scales2")`

---


```{js}
d3.select("body").append("svg:svg") b .data([[1, 1.2, 1.7, 1.5, .7]]) .attr("width", 150)
.attr("height", 150)
  .selectAll("path")
    .data(d3.layout.pie())
  .enter().append("svg:path")
    .attr("transform", "translate(75,75)")
    .attr("d", d3.svg.arc().outerRadius(70));
```



```{css, eval = TRUE, echo = F}
.remark-code{line-height: 1.5; font-size: 65%}
```
