---
title: "Introduction to the Bootstrap"
author: "Gina Reynolds, June 2019"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", "ninjutsu"]
    nature:
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
knitr::opts_chunk$set(fig.height = 6, out.width = "100%", comment = " ", cache = F)
source("https://raw.githubusercontent.com/EvaMaeRey/little_flipbooks_library/master/xaringan_reveal_parenthetical.R")
```



---

# Introduction

This is a minimal example to demonstrate how to create a flipbook with data from #TidyTuesday.  It walks through data wrangling and plots pipelines made with the Tidyverse.  The functions that make this possible are the work of Emi Tanaka, Garrick Aden-Buie and myself, and are built for Xaringan, an Rmarkdown file type for creating presentation slides; the functions make use of the function `knitr:::knit_code$get()`.

The code to create the flipbook is an .Rmd that you can download [**here**](https://raw.githubusercontent.com/EvaMaeRey/little_flipbooks_library/master/tidytuesday_minimal_example/tidytuesday_minimal_example.Rmd).

---

Interested in more flipbooks? Check out 

- [the ggplot flipbook](https://evamaerey.github.io/ggplot_flipbook/ggplot_flipbook_xaringan.html)
- [The Tidyverse in Action](https://evamaerey.github.io/tidyverse_in_action/tidyverse_in_action.html)




For more about Xaringan:

- [Xaringan presentation slides](https://slides.yihui.name/xaringan/)

The sequential workflow of the Tidyverse makes incremental display of pipelines and ggplot statements ideal:  

- [www.tidyverse.org](https://www.tidyverse.org/)


---

---

# Set up

Okay. Let's load the the `reveal for xaringan` functions for "flipbooking" and the `tidyverse`.

```{r}
source(file = "https://raw.githubusercontent.com/EvaMaeRey/little_flipbooks_library/master/xaringan_reveal_parenthetical.R")
```

And load the tidyverse.

```{r}
library(tidyverse)
```

---

```{r sample, echo = F}
set.seed(2019)
tibble(x = rnorm(30)) %>% 
  mutate(y = x * .4 + rnorm(30)) %>% 
  mutate(id = 1:n()) ->
my_data
ggplot(data = my_data) +
  aes(x = x) +
  aes(y = y) +
  geom_point() +
  geom_smooth(method = lm, se = F)
```

`r apply_reveal("sample")`

---

```{r bootsamples, echo = F}
my_data %>% 
  crossing(boot_sample_id = 1:100) %>% 
  group_by(boot_sample_id) %>% 
  sample_frac(size = 1, 
              replace = TRUE) %>% 
  group_by(id, boot_sample_id) %>%
  mutate(times_sampled = n()) ->
boot_samples
```


`r apply_reveal("bootsamples")`

---

# Ensemble plot



```{r plotboot, echo = F}
boot_samples %>% 
  filter(boot_sample_id <= 12) %>% 
  ggplot() +
  aes(x = x) +
  aes(y = y) +
  geom_point(size = 3, col = "magenta") +
  facet_wrap(~ paste("boot sample", boot_sample_id)) +
  aes(label = paste0(times_sampled, "X")) +
  geom_text(size = 2, col = "grey", aes(alpha = NULL)) +
  aes(group = boot_sample_id) +
  aes(alpha = times_sampled) +
  geom_smooth(method = lm, se = F) +
  ggdark::dark_mode() +
  theme(legend.position = "none")
```

---

`r apply_reveal("plotboot")`

---

```{r}
boot_model <- function(df){
  lm(y ~ x, data = df)
}
```


```{r modelboot, echo = F}
boot_samples %>%
  group_by(boot_sample_id) %>% 
  nest() %>% 
  mutate(model = map(data, boot_model)) %>%   # model results are summarised in tidy dataframes using broom
  mutate(glance  = map(model, broom::glance),   
         augment = map(model, broom::augment),  
         tidy    = map(model, broom::tidy)) ->   
boot_samples_models
```

---

`r apply_reveal("modelboot")`

---

# Slope coefficient confidence interval


```{r confint, echo = F}
boot_samples_models %>% 
  unnest(tidy) %>% 
  filter(term == "x") ->
x_estimates  
ggplot(data = x_estimates) +
  aes(x = estimate) +
  geom_histogram() +
  geom_rug() +
  geom_vline(xintercept = 
               quantile(x_estimates$estimate, 
                        probs = c(.05, .95)), 
             col = "red",
             linetype = "dashed")
```

---

`r apply_reveal("confint")`

---

```{r confintplus, echo = F}
boot_samples_models %>% 
  unnest(tidy) %>% 
  select(boot_sample_id, estimate, term) %>% 
  spread(key = term, value = estimate) %>% 
  ggplot() +
  aes(x = x) +
  aes(y = `(Intercept)`) +
  geom_point() +
  geom_density_2d() 
```


`r apply_reveal("confintplus")`

---


```{css, eval = TRUE, echo = F}
.remark-code{line-height: 1.5; font-size: 80%}
```

```{r xaringan-themer, include = FALSE}
library(xaringanthemer)


xaringanthemer::write_xaringan_theme(text_font_size = ".75cm",
                                     text_color = "darkblue")

xaringanthemer::mono_light(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "200", "200i"),
  code_font_google   = google_font("Droid Mono"),
  text_font_size = ".6cm",
  code_font_size = ".3cm")
  
```

