---
title: "Proposed code parsing for flipbooks"
author: "Gina Reynolds, May 2019"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["kunoichi", "ninjutsu"]
    nature:
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r setup, include=F}
knitr::opts_chunk$set(fig.width = 6, message = F, warning = F, 
                      comment = "", cache = F)
```




---

# Introduction

This is a minimal example to demonstrate how to create a flipbook with data from #TidyTuesday.  It walks through data wrangling and plots pipelines made with the Tidyverse.  The functions that make this possible are the work of Emi Tanaka, Garrick Aden-Buie and myself, and are built for Xaringan, an Rmarkdown file type for creating presentation slides; the functions make use of the function `knitr:::knit_code$get()`.

The code to create the flipbook is an .Rmd that you can download [**here**](https://raw.githubusercontent.com/EvaMaeRey/little_flipbooks_library/master/tidytuesday_minimal_example/tidytuesday_minimal_example.Rmd).

---

Interested in more flipbooks? Check out 

- [the ggplot flipbook](https://evamaerey.github.io/ggplot_flipbook/ggplot_flipbook_xaringan.html)
- [The Tidyverse in Action](https://evamaerey.github.io/tidyverse_in_action/tidyverse_in_action.html)

For more about Xaringan:

- [Xaringan presentation slides](https://slides.yihui.name/xaringan/)

The sequential workflow of the Tidyverse makes incremental display of pipelines and ggplot statements ideal:  

- [www.tidyverse.org](https://www.tidyverse.org/)


---

# Set up

Okay. Let's load the the `reveal for xaringan` functions for "flipbooking" and the `tidyverse`.

```{r}
source(file = "../xaringan_reveal_parentheses_balanced.R")
```

And load the tidyverse.

```{r}
library(tidyverse)
```

And load the data from the tidytuesday github page. 

```{r}
library(gapminder)
library(tidyverse)
```


---

# Suppose we have some code

```{r the_code, eval = F}
gapminder %>%             # the data 
  filter(year == (2000 + 7)) %>%  # subset 
  ggplot() +              # pipe to ggplot 
  aes(x = gdpPercap) +
  aes(y = lifeExp) +  
  geom_point() +
  aes(color = 
  paste("continent", 
        continent) 
          ) ->  
my_plot 
```

---

`r apply_reveal("the_code")`


---

```{r}
parse_code(knitr:::knit_code$get("the_code"))
```


# proposed parsing steps

```{r parse}
tibble(code = 
         knitr:::knit_code$get("the_code")) %>% 
  mutate(user_reveal = 
           str_detect(code, "#REVEAL")) %>% # Handle user defined pause points
  mutate(code = 
           str_remove(code, "#REVEAL")) %>% # pull out any comments
  separate(col = code,                   
           into = c("code", "comment"), 
           sep = "#") %>% 
  mutate(comment = 
           str_trim(comment)) %>% 
  mutate(comment = 
           paste0("  # ", comment)) %>%  
  mutate(comment = 
           ifelse("  # NA" == comment, 
                  "", comment)) %>% 
  mutate(code = 
           str_remove(code, "\\s+$")) %>% 
  mutate(connector = 
           str_extract(code, "%>%$|\\+$|->$")) %>% 
  mutate(connector = replace_na(connector, "")) %>% 
  mutate(code = 
           str_remove(code, "%>%$|\\+$|->$")) %>% 
  mutate(num_open_par =                       
           str_count(code, "\\(|\\{|\\[")) %>% # Counting open parentheses
  mutate(num_closed_par = 
           str_count(code, "\\)|\\}|\\]")) %>%  # Counting closed parentheses
  mutate(balanced_par = (cumsum(num_open_par) - 
                           cumsum(num_closed_par)) == 0) ->  # Check for balance
parsed
```

---

```{r}
as_data_frame(parsed)
```



---

# Now build to different points: 

```{r build}
4 -> up_to
parsed %>% 
  mutate(reveal = 1:n() <= up_to) %>% 
  mutate(out = paste0(code, "", connector, "", comment)) %>% 
  filter(reveal) %>% 
  select(out) ->
up_to_result
paste(up_to_result$out, collapse = "\\n") %>% 
  str_remove("%>%$|\\+$|->$") %>% 
  cat()
```  


---






```{css, eval = TRUE, echo = F}
.remark-code{line-height: 1.5; font-size: 30%}
```


