---
title: "UFOs"
subtitle: "Joel Soroos | made with Xaringan"
author: "<br><br>Edited by Gina Reynolds, 2019"
date: "<br>"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["kunoichi", "ninjutsu"]
    nature:
      ratio: "16:9"
      highlightLines: true
---

```{r, include=FALSE}
knitr::opts_chunk$set(fig.height = 6, out.width = "100%", comment = " ", cache = T, echo = TRUE, warning = F, message = F)
source("https://raw.githubusercontent.com/EvaMaeRey/little_flipbooks_library/master/xaringan_reveal_parentheses_balanced.R")
```


# UFO sitings

This work is by Joel Soros who focuses on UFO sitings in North Carolina.  


---

## Set up



```{r, warning = F, results = FALSE, message = FALSE}
library(tidyverse)
url <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-06-25/ufo_sightings.csv"
ufo_raw <- readr::read_csv(file = url) 
```

---

## First a little data cleaning

---

```{r transform_ufo_nc, message = F, echo = F, eval = T}
ufo_raw %>%
      janitor::clean_names() %>% 
      select(date_time, city_area, 
             state, latitude, 
             longitude, encounter_length) %>%
  rename(lat = latitude,
         long = longitude) %>% 
      filter(
         state == "nc",
         lat > 30,      # remove borders erroneously listed as NC outside of state borders
         lat < 37,      # remove borders erroneously listed as NC outside of state borders
         long < -75     # remove borders erroneously listed as NC outside of state borders
         ) %>%
      mutate(
         encounter_length = encounter_length/3600,              #convert seconds to hours
         date_time = as.Date(date_time, format = "%m/%d/%Y")
         ) ->
ufo  
```


---

`r apply_reveal("transform_ufo_nc")`


---

## prepping map data

He also uses ggplot2's map_data to plot the North Carolina polygon.  


```{r}
map_borders <- 
  ggplot2::map_data("state", 
           region = "north carolina") 
```

---

# The basic plot

---

```{r build_ufo_nc, echo = F, eval = T}
ufo %>% 
  ggplot() +
  aes(x = long) +
  aes(y = lat) +
  # plot nc borders
  geom_polygon(data = map_borders,
               mapping = aes(group = group),
               color = "black", 
               fill = "grey25", 
               size = 1.15) +
  # plot UFO encounters
  aes(size = encounter_length) +
  geom_point(color = "green", aes(group = NULL)) +
  coord_fixed(1.3) +
  scale_size_continuous(breaks = c(1, 10, 100)) +
  labs(title = "UFOs over North Carolina\n") +
  labs(size = "Encounter (hrs)") +
  labs(caption = "\nEach dot represents a reported UFO sighting between 1995 and 2014.  \nSource: National UFO Reporting Center  | Visualization: Joel Soroos @soroosj") ->
basic_plot
```

`r apply_reveal("build_ufo_nc")`


---  
  
## Annotation layer

---

```{r annotation_ufo, echo = F, eval = T}
basic_plot +
  annotate(geom = "text",
        label = "30 hour encounter\nin Deep Gap (2009)",
        size = 3, hjust = 0, color = "green", 
        family = "Rockwell",
        x = -84.9, y = 36.1
        ) +
  annotate(geom = "curve", 
           x = -83.2, y = 36.2, 
           xend = -81.7, yend = 36.27,
        arrow = arrow(length = unit(0.2, "cm")),
        size = 0.4, color = "green", curvature = -0.4
        ) +
  # Gastonia encounter annotation
     annotate("text",
        label = "120 hour encounter\nin Gastonia in 1993.",
        size = 3, hjust = 0, color = "green", 
        family = "Rockwell",
        x = -83.1, y = 34.5
        ) +
     annotate(geom = "curve",
              x = -82.27, y = 34.65, 
              xend = -81.37, yend = 35.2,
        arrow = arrow(length = unit(0.2, "cm")), 
        size = 0.4, color = "green", curvature = -0.3
        ) ->
ufo_plot
```

`r apply_reveal("annotation_ufo")`


---

# Adjusting themes

---

```{r ufo_theme, echo = F, eval = F}
ufo_plot +
  ggdark::dark_mode(.theme = 
                      theme_minimal()) +
  theme(axis.title = element_blank()) +
  theme(axis.text = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(text = element_text(family = "Rockwell", 
                            color = "green")) +
  theme(plot.title = element_text(hjust = 0.5, 
                                  size = 18)) +
  theme(plot.caption = element_text(hjust = 0, 
                                    size = 8)) +
  theme(legend.title = element_text(size = 10, 
                                    hjust = 0.5, 
                                    vjust = 0.5)) +
  theme(legend.text = element_text(size = 9, 
                                   hjust = 0.5, 
                                   vjust = 0.5)) +
  theme(legend.position = c(0.82, 0.18)) +
  theme(legend.justification = c(0, 1)) +
  theme(legend.key.size = unit(0.1, 'lines'))
```

---



`r apply_reveal("ufo_theme")`

---

```{r, include = F}
save_complete_plot_from_chunk(chunk_name = "ufo_theme")
```


```{css, eval = TRUE, echo = F}
.remark-code{line-height: 1.5; font-size: 40%}
```