---
title: "Racing Barchart with gganimate"
subtitle: "a quick flipbook | made with Xaringan"
author: "<br><br>Gina Reynolds, April 2019"
date: "<br>"
output:
  xaringan::moon_reader:
    chakra: libs/remark-latest.min.js
    lib_dir: libs
    css: ["kunoichi", "ninjutsu"]
    nature:
      ratio: "16:9"
      highlightLines: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.height = 6, out.width = "100%", comment = " ", cache = T, message = F, warning = F, dpi = 300)

source("https://raw.githubusercontent.com/EvaMaeRey/little_flipbooks_library/master/xaringan_reveal_parenthetical.R")
```


# Introduction

This book walks through data prep and plotting a racing barchart with ggplot and gganimate. 



---

# Set up


```{r}
library(gapminder)
library(tidyverse)
```


```{r prep_data, echo = F, eval = T}
gapminder %>% 
  select(country, pop, year, continent) %>% 
  group_by(year) %>% # for each year we assign a rank
  arrange(year, -pop) %>% 
  mutate(rank = 1:n()) %>% # assign ranking
  filter(rank <= 10) ->
ranked_by_year
```

---

`r apply_reveal("prep_data")`

---



```{r, echo = T}
my_theme <- theme_classic(base_family = "Times") +
  theme(axis.text.y = element_blank()) +
  theme(axis.ticks.y = element_blank()) +
  theme(axis.line.y = element_blank()) +
  theme(legend.background = element_rect(fill = "linen")) +
  theme(plot.background = element_rect(fill = "linen")) +
  theme(panel.background = element_rect(fill = "linen"))
```
  
---
# Static plot
  
```{r static_bar, eval = T, echo = F}  
ggplot(data = ranked_by_year) +
  aes(group = country, fill = continent) +
  aes(xmin = 0 , 
      xmax = pop / 1000000) +
  aes(ymin = rank - .45, 
      ymax = rank + .45) +
  scale_y_reverse() +
  scale_x_continuous(
    limits = c(-300, 1400),
    breaks = c(0, 400, 800, 1200),
    labels = c(0, 400, 800, 1200)) +
  labs(fill = "") +
  geom_rect(alpha = .7) +
  labs(x = 'Population (millions)') +
  aes(label = country, y = rank) +
  geom_text(col = "gray13", 
            hjust = "right", 
            x = -50) + 
  labs(y = "") +
  scale_fill_viridis_d(option = "magma", 
                       direction = -1) +
  geom_text(x = 1000 , y = -10, 
            family = "Times", 
            aes(label = as.character(year)), 
            size = 30, col = "grey18") +
  my_theme ->
g
```

---
`r apply_reveal("static_bar")`

---

# Now set up the animation

Next we add the animation statement, which uses a function from the gganimate package. 

```{r, eval = T}
library(gganimate) 
options(gganimate.nframes = 20)
```

---

```{r ani_final, eval = F, echo = F, fig.height=5, fig.width=4}
g + gganimate::transition_time(year)
```

`r apply_reveal("ani_final")`


---

# A lessons learned about racing barcharts.

My first implementation was with GDP per capita.  This resulted in less visually pleasing because of rank changing beyond adjacent neighbors between time periods.  The version of the data used is only every five years, so such dramatic switching between time periods isn't surprising.  Using population gave a better result. 


```{css, eval = TRUE, echo = F}
.remark-code{line-height: 1.5; font-size: 60%}
```










