---
title: "A Minimal Flipbook"
author: "Gina Reynolds, May 2019"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["kunoichi", "ninjutsu"]
    nature:
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r setup, include=F}
knitr::opts_chunk$set(fig.width = 6, message = F, warning = F, 
                      comment = "", cache = T)
options(tibble.print_max = Inf) 
```




---

# Introduction

This is a minimal example to demonstrate how to create a flipbook that will walk through data wrangling and plots pipelines made with the Tidyverse.  The functions that make this possible are the work of Emi Tanaka, Garrick Aden-Buie and myself, and are built for Xaringan, an Rmarkdown file type for creating presentation slides; the functions make use of the function `knitr:::knit_code$get()`.

Interested in more flipbooks? Check out 

- [the ggplot flipbook](https://evamaerey.github.io/ggplot_flipbook/ggplot_flipbook_xaringan.html)
- [The Tidyverse in Action](https://evamaerey.github.io/tidyverse_in_action/tidyverse_in_action.html)

For more about Xaringan:

- [Xaringan presentation slides](https://slides.yihui.name/xaringan/)

The sequentail workflow of the Tidyverse makes incremental display of pipelines ideal:  

- [www.tidyverse.org](https://www.tidyverse.org/)


---

# Set up

Okay. Let's load the the `reveal for xaringan` functions for "flipbooking" and the `tidyverse`.

```{r}
source(file = "https://raw.githubusercontent.com/EvaMaeRey/little_flipbooks_library/master/xaringan_reveal_parenthetical.R")
library(tidyverse)
```



---

# Where we are going: 

We'll create this checkerboard (something Andrew Heiss was recently working on).  I have **echo** set to FALSE in the code chunk options here so that you don't see the code, and **eval** to TRUE so that the plot output is produced.  The code chunk is given the name "checkerboard", and this is used in the in apply_reveal function, which will break down the code by wrangling and plot statements. 

```{r checkerboard, echo = F, eval = T, fig.height=5}
crossing(x = 1:8, y = 1:8) %>%
  mutate(value = (x + y) %% 2) %>% 
  ggplot() + 
  aes(x = x, 
      y = y, 
      fill = factor(value)) +
  geom_tile() +
  theme_void() +
  scale_fill_manual(
    values = c("thistle", 
               "darkgrey"), 
    guide = F) + 
  coord_equal()
```
 

 
---

# How do we get there?

In the next slide, we'll walk through the code that produces this plot, and the output along the way. We use the code **apply_reveal("checkerboard")** in-line to access the code from the code chunk called *checkerboard*.
 
---

`r apply_reveal("checkerboard")`




```{css, eval = TRUE, echo = F}
.remark-code{line-height: 1.5; font-size: 100%}
```


