---
title: "ols"
subtitle: "made with flipbookr and xaringan"
author: "Gina Reynolds, Jan 2020"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", ninjutsu]
    nature:
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---




```{r xaringan_look, include = F}
xaringanthemer::mono_accent(
  code_highlight_color = "rgba(255, 211, 0, .35)",
  base_color = "#03418A")
knitr::opts_chunk$set(warning = F, message = F)
```

```{r, include = F}
library(tidyverse)
library(flipbookr)
```

---

`r flipbookr::chunk_reveal("basic", break_type = 1)`


```{r basic, include = F}
cars %>% 
  ggplot() +
  labs(title = "Fitting an OLS model to data in ggplot2") +
  aes(x = speed) +
  aes(y = dist) +
  geom_point(color = "steelblue",
             size = 2) +
  geom_smooth(method = lm, se = F)
```

---

OLS in base R


---

`r flipbookr::chunk_reveal("base_ols")`

```{r base_ols, include = F}
# fitting a model
lm(formula = dist ~ speed,
   data = cars) ->
cars_model
# using summary() to
# to pull out more info
cars_model %>% 
  summary() %>% 
  .$residuals
```

---

# But {broom} is cool ...

--

...anticipating that we might want lots of this info pulled out systematically as data frames


--

Broom returns dataframes at 3 levels

- broom::glance(), overall model statistics
- broom::tidy(), covariate level model statistics
- broom::augment(), observation level statistics

---

# *Model* level stats

```{r}
cars_model %>% 
  broom::glance()
```


---

# *Covariate* level stats


```{r}
cars_model %>% 
  broom::tidy()
```

---

# *Observation* level


```{r}
cars_model %>% 
  broom::augment() 
```

---

`broom::augment()` is good for plotting those residuals and fitted values, and checking model assumptions.

---

`r flipbookr::chunk_reveal("augment")`


```{r augment, include = F}
cars_model %>% 
  broom::augment() %>% 
  ggplot() +
  labs(title = "Plot of model fit") +
  aes(x = speed) +
  aes(y = dist) +
  geom_point(col = "steelblue") +
  geom_smooth(method = lm, se = F) +
  geom_point(aes(y = .fitted)) +
  aes(xend = speed) +
  aes(yend = .fitted) +
  geom_segment(color = "red", 
               linetype = "dashed") +
  # input value of 21 miles per hour
  geom_vline(xintercept = 21,
             linetype = "dotted") +
  # yields predicted for y, stopping distance
  geom_hline(yintercept = 
               predict(cars_model, 
                       data.frame(speed = 21)),
             linetype = "dotted")
```

---

Checking assumption:

residuals not correlated with predictions...

---

`r flipbookr::chunk_reveal("residuals_plot")`


```{r residuals_plot, include = F}
cars_model %>% 
  broom::augment() %>%
  ggplot() +
  labs(title = "Residuals v. Fitted") +
  aes(x = .fitted) +
  aes(y = .resid) +
  geom_point(col = "steelblue") +
  geom_hline(yintercept = 0,
             lty = "dashed") +
  geom_smooth(se = F, span = 1, 
              color = "red")
```


---

# Checking for outliers with Cook's distance


---

`r flipbookr::chunk_reveal("cooks")`

```{r cooks, include = F}
cars_model %>% 
  broom::augment() %>%
  mutate(id = row_number()) %>% 
  ggplot() +
  labs(title = "Cook's Distance, by ID (row number)") +
  labs(subtitle = "Cook's Distance, a measure of influence via a leave one out procedure") +
  aes(x = id) +
  aes(y = .cooksd) +
  geom_point() +
  aes(label = id) +
  geom_text(check_overlap = T, 
            nudge_y = .015)
```

---


```{r, echo = F, eval = F}
cars_model %>% 
  broom::augment() %>%
  mutate(id = row_number()) %>% 
  ggplot() +
  labs(title = "Cook's Distance v. standardized residual") +
  labs(subtitle = "Cook's Distance, a measure of influence via a leave one out procedure") +
  aes(x = id) +
  labs(x = "id (dataframe row number)") +
  aes(y = .cooksd) +
  geom_point() +
  aes(label = id) +
  geom_text(check_overlap = T, 
            nudge_y = .015)
```

---

Are the residuals normally distributed? 


---

`r flipbookr::chunk_reveal("qqplot")`

```{r qqplot, include = F}
cars_model %>% 
  broom::augment() %>% 
  mutate(expected = qnorm((1:n() - .5)/n())) %>% 
  ggplot() +
  labs(title = "Normal q-q plot") +
  aes(y = sort(.std.resid)) +
  aes(x = expected) +
  geom_rug() +
  geom_point() +
  coord_equal(ratio = 1, 
              xlim = c(-4, 4),
              ylim = c(-4, 4)) +
  # add line of equivilance
  geom_abline(intercept = 0, 
              slope = 1, 
              linetype = "dotted")
```


---

# 


```{r}
cars_model
```




---



```{r, eval = F, echo = F}
cars_model %>% 
  plot()
```

---

# multiple regression

--
and multiple models in a single table

---

# subsetting to 2002

```{r}
library(gapminder)
gapminder %>% 
  filter(year == 2002) ->
gapminder_2002
```

---

`r flipbookr::chunk_reveal("multi_model")`


```{r multi_model, include = F}
lm(lifeExp ~ gdpPercap, 
   data = gapminder_2002) ->
model1
lm(lifeExp ~ gdpPercap + pop, 
   data = gapminder_2002) ->
model2
lm(lifeExp ~ gdpPercap + 
     continent, 
   data = gapminder_2002) ->
model3
```


---

```{r, results = "asis"}
stargazer::stargazer(
  model1, model2, model3,
  # if you want to work interactivevly, change "html" to "text" 
  # note, chunk results option is "asis"
  type = "html",  
  dep.var.labels = "Life Expectancy (years)", 
  covariate.labels = c("GDP per cap (Thousands $US)", 
                       "Population (Millions)"), 
  style = "qje", 
  font.size = "small",
  title = "Models of Life Expectancy in 2002")
```





```{css, eval = TRUE, echo = F}
.remark-code{line-height: 1.5; font-size: 80%}
```
