---
title: "Flipbooks"
author: Evangeline Reynolds, Korbel School of International Studies
date: August 8, 2019
output:
  html_document:
    df_print: paged
  html_notebook:
    nature:
      beforeInit: https://platform.twitter.com/widgets.js
---




```{r, include=F}
knitr::opts_chunk$set(cache = T, echo = F, warning = F, message = F, cache = T, error = F)
```


## What are flipbooks?  

"Flipbooks" are a new tool that present side-by-side, aligned, incremental code-output evolution via automated code parsing and reconstruction.

Visual superimposition of new code on previous code and new output on previous output yields a ["blink comparator"](https://en.wikipedia.org/wiki/Blink_comparator){target="_blank"} type experience, and incrementalism assists viewers in linking specific components of code with resultant behavior.  Flipbooks are promising as communication, educational, and reference medium.

Click on the example below. Then you should be able to "flip" (left and right arrows) through the book to see how to make an animated racing bar chart of population size in the most populous countries over time.   

```{r}
knitr::include_url("https://evamaerey.github.io/little_flipbooks_library/racing_bars/racing_barcharts.html#1")
```


## Origins

In mid 2018 Garrick Aden-Buie, a data scientist at the Moffitt Cancer Center, created and debuted a side-by-side format for code and figure output in the slide-show tool, Xaringan, an RMarkdown tool.  He demonstrated in incremental steps how someone might build up a plot using the popular R Stats package ggplot2.  In August 2018, I emulated this work, but with an entirely sequential and more incremental workflow that I call "slow ggplotting".  In September, Emi Tanaka, a Lecturer in Statistics at the University of Sydney, implemented a similar sequential workflow demonstration, that was partially automated.  

I approached Garrick and Emi in October about working together to create a "flipbook" of exemplary data visualizations.  We soon had a fully automated framework to parse code, and present its reconstruction along with associated output!  

The code designed to parse and reconstruct code was then tested in small projects and refined.  Then I created a flipbook of a sizable collection of plots that I had coded over about about a year and a half in learning ggplot.  After rewriting this code in a "slow" ggplot style, and providing written description to introduce the plots, I publicized the work in February, 2019; the flipbook was met an enthusiastic reception.  

Since then, a number of changes have been made to the foundational code to improve the tool and to enhance its usefulness. 



<!-- https://www.garrickadenbuie.com/blog/decouple-code-and-output-in-xaringan-slides/ -->
<!-- https://emitanaka.rbind.io/post/knitr-knitr-code/ -->

<!-- - the code automating the code-walk through has been updated to "pause" at more "connectors" in the code -->
<!-- - to allow code wrapping -->
<!-- - allow the user to indicate at what point walk-through should be "paused" -->
<!-- - adding balanced parentheses as a condition of   -->

 <!-- The usability of flipbooks is also being improved by creating tools for the creation of visual table of contents, from which users can link to specific visualizations to view code-plot evolution.    -->

<!-- ## Origins - the ideas hose -->

<!-- <blockquote class="twitter-tweet"><p lang="en" dir="ltr">Good ideas! Very <a href="https://twitter.com/hashtag/ggplot?src=hash&amp;ref_src=twsrc%5Etfw">#ggplot</a>. A layered presentation of graphics. <a href="https://t.co/M4tRB9bGS5">https://t.co/M4tRB9bGS5</a></p>&mdash; Gina Reynolds (@EvaMaeRey) <a href="https://twitter.com/EvaMaeRey/status/991269331257487366?ref_src=twsrc%5Etfw">May 1, 2018</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script> -->

<!-- <blockquote class="twitter-tweet"><p lang="en" dir="ltr">I like this implementation even better. I&#39;m using aes() outside of the ggplot() function or a geom function seems not to be conventional. I just tried and see that it works in this case as I was hoping. But, should it be avoided? bad style? <a href="https://twitter.com/StatGarrett?ref_src=twsrc%5Etfw">@StatGarrett</a> <a href="https://t.co/GxUGo1agHU">pic.twitter.com/GxUGo1agHU</a></p>&mdash; Gina Reynolds (@EvaMaeRey) <a href="https://twitter.com/EvaMaeRey/status/992341968125493248?ref_src=twsrc%5Etfw">May 4, 2018</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script> -->


<!-- <blockquote class="twitter-tweet"><p lang="en" dir="ltr">Really dig how <a href="https://twitter.com/grrrck?ref_src=twsrc%5Etfw">@grrrck</a> builds up <a href="https://twitter.com/hashtag/ggplot2?src=hash&amp;ref_src=twsrc%5Etfw">#ggplot2</a> syntax w/ 📊:<br>📽 &quot;A Gentle Guide to the Grammar of Graphics with ggplot2&quot; <a href="https://t.co/2Okhri7Hox">https://t.co/2Okhri7Hox</a> <a href="https://twitter.com/hashtag/rstats?src=hash&amp;ref_src=twsrc%5Etfw">#rstats</a> <a href="https://twitter.com/hashtag/dataviz?src=hash&amp;ref_src=twsrc%5Etfw">#dataviz</a> <a href="https://t.co/qO43IeTtHG">pic.twitter.com/qO43IeTtHG</a></p>&mdash; Mara Averick (@dataandme) <a href="https://twitter.com/dataandme/status/995650384155693061?ref_src=twsrc%5Etfw">May 13, 2018</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script> -->

<!-- https://pkg.garrickadenbuie.com/gentle-ggplot2/#46 -->

<!-- <blockquote class="twitter-tweet"><p lang="en" dir="ltr">Here, building up a <a href="https://twitter.com/hashtag/ggplot2?src=hash&amp;ref_src=twsrc%5Etfw">#ggplot2</a> as slowly as possible, <a href="https://twitter.com/hashtag/rstats?src=hash&amp;ref_src=twsrc%5Etfw">#rstats</a>. Incremental adjustments. <a href="https://twitter.com/hashtag/rstatsteachingideas?src=hash&amp;ref_src=twsrc%5Etfw">#rstatsteachingideas</a> <a href="https://t.co/nUulQl8bPh">pic.twitter.com/nUulQl8bPh</a></p>&mdash; Gina Reynolds (@EvaMaeRey) <a href="https://twitter.com/EvaMaeRey/status/1029104656763572226?ref_src=twsrc%5Etfw">August 13, 2018</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script> -->

<!-- <blockquote class="twitter-tweet"><p lang="en" dir="ltr">Inspired by <a href="https://twitter.com/grrrck?ref_src=twsrc%5Etfw">@grrrck</a> and <a href="https://twitter.com/EvaMaeRey?ref_src=twsrc%5Etfw">@EvaMaeRey</a>, made the kunoichi + ninjutsu (ninja-theme) version of <a href="https://twitter.com/hashtag/ggplot?src=hash&amp;ref_src=twsrc%5Etfw">#ggplot</a> tutorial although Garrick already does explaining this in his excellent blog <a href="https://t.co/msXfg14Ztn">https://t.co/msXfg14Ztn</a>. Gist for ninja-theme here: <a href="https://t.co/soHH4Qvz4F">https://t.co/soHH4Qvz4F</a> <a href="https://twitter.com/hashtag/rstats?src=hash&amp;ref_src=twsrc%5Etfw">#rstats</a> <a href="https://t.co/YlRHAGnaUm">pic.twitter.com/YlRHAGnaUm</a></p>&mdash; Emi Tanaka 🌾 (@statsgen) <a href="https://twitter.com/statsgen/status/1041279648452108289?ref_src=twsrc%5Etfw">September 16, 2018</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script> -->




## Use cases

### Code/output demonstration

The big debut of the flipbook technique was unveiling in its application to a substantial collection of data visualization code: [the ggplot flipbook](https://evamaerey.github.io/ggplot_flipbook/ggplot_flipbook_xaringan.html){target="_blank"} in February of 2019.  Shortly thereafter, the reveal code was extended to handle data manipulation pipelines which allowed for the creation of [The Tidyverse in Action](https://evamaerey.github.io/tidyverse_in_action/tidyverse_in_action.html){target="_blank"}.   

Since then flipbook tool has been applied in explaining how to create visualizations important to statistical pedagogy, ([The Galton Board flipbook](https://evamaerey.github.io/little_flipbooks_library/galton_board/galton_board.html){target="_blank"}
); political decision rules [D’Hondt/Jefferson Allocation](https://evamaerey.github.io/little_flipbooks_library/dhondt_jefferson_allocation/dhondt_jefferson_allocation){target="_blank"}; and custom plot types like [a racing barchart](https://evamaerey.github.io/little_flipbooks_library/racing_bars/racing_barcharts.html){target="_blank"}.

Returning to the initial focus of data visualization using ggplot, new resources have been created focusing on more providing systematic and theoretical foundations for understanding ggplot2 -- the implementation of the powerful grammar of graphics framework in R -- including [a ggplot grammar guide](https://evamaerey.github.io/ggplot2_grammar_guide/ggplot2_grammar_guide#1){target="_blank"} and [ggplot themes](https://evamaerey.github.io/little_flipbooks_library/taming_themes_in_ggplot/taming_ggplot_themes.html){target="_blank"}.  

Use of the framework has also been explored outside of its ["Tidyverse"](https://www.tidyverse.org/){target="_blank"} roots in [data.table](https://evamaerey.github.io/little_flipbooks_library/data.table/data.table){target="_blank"}, and [leaflet's interactive maps for R](https://evamaerey.github.io/little_flipbooks_library/leaflet/leaflet#1){target="_blank"} for example.  

### Use in traditional presentations

An additional adaptation of the parsing and reveal code allows users to slowly reveal the components of a plot without also featuring the code to produce the plot in [minimal example #2](https://evamaerey.github.io/little_flipbooks_library/minimal_example/minimal_example){target="_blank"}, a strategy that Hans Rosling used in his stunning presentations, and a technique that is recommended by data journalist Mona Chalabi and Harvard  Professor of Government Matthew Blackwell. 


## What's next

### Supporting new user adoption: examples, templates, and R package

A code repository for the smaller flipbooks projects "lives" in one place, [here](https://github.com/EvaMaeRey/little_flipbooks_library), for reference.   Larger projects' code is also open access and is available for examination and emulation.

The next frontier is making the technique easy to use.  One step to this end was to use others' code to create Flipbooks as was done in [#TidyTuesday walk through](https://evamaerey.github.io/tidytuesday_walk_through/tidytuesday_highlights.html){target="_blank"}.  Four visualizer's enerously allowed me to use their beautiful work.  Problems with the code parsing was indeed discovered and addressed through this process!

Additionally, I'm working on and testing materials to help people create their own books with minimal hassle.  A guide exists in the [minimal example #1](https://evamaerey.github.io/little_flipbooks_library/tidytuesday_minimal_example/tidytuesday_minimal_example){target="_blank"} and the [checkerboard](https://evamaerey.github.io/little_flipbooks_library/checkerboard/checkerboard.html){target="_blank"}.  In a June 2019 workshop participants at the University of Illinois, successfully produced their own flipbook using the minimal example as a guide.  

Given the third party success with the current code, we're confident that an R package with the flipbook functions can be useful to a wide audience.  

### Theorticeal description

I'm currently working on a manuscript thinking about flipbooks from a theoretical perspective including a discussion of humnition, describing the format and our particular implementation, as well as articulating its merits. 

### Extension to non sequential code

Sequentially revealing code is the current implementation of flipbooks. However, there are certainly circumstances in which it would make more sense to reveal things out of order to better communicate about code input and subsequent output behavior.  Tackling non sequential reveal is therefore on the to do list.   

## Reception

The feedback on the flipbook format has been tremendously encouraging, and a big motivation to keep working on this project.  The flipbook was a [top story on Hacker News](https://news.ycombinator.com/item?id=19356354) and was featured as a highlight in the [R Weekly](https://www.r-craft.org/r-news/r-weekly-2019-07-flipbook-package-building/) newsletter.


Some of the most encouraging feedback about flipbooks has come from individuals, via Twitter.  Their comments can be viewed [here](https://evamaerey.github.io/little_flipbooks_library/about/cheers_for_flipbooks).

Feedback comes from individuals around the world including the following locations!

```{r, results='asis'}
library(tidyverse)
load("data_products/tweets.Rdata")
df %>% 
  select(location) %>% 
  drop_na() %>% 
  pull(location) %>% 
  cat(sep = " | ")
```



<!-- ```{r, out.width="48%"} -->
<!-- knitr::include_url("https://evamaerey.github.io/ggplot_flipbook/ggplot_flipbook_xaringan.html") -->
<!-- knitr::include_url("https://evamaerey.github.io/tidyverse_in_action/tidyverse_in_action.html") -->
<!-- knitr::include_url("https://evamaerey.github.io/ggplot2_grammar_guide/ggplot2_grammar_guide") -->
<!-- knitr::include_url("https://evamaerey.github.io/tidytuesday_walk_through/tidytuesday_highlights.html") -->
<!-- ``` -->

