---
title: "Flipbooks: Presenting code and its output side-by-side and step-by-step"
author: Evangeline Reynolds, Korbel School of International Studies
date: August 8, 2019
output:
  html_document:
    df_print: paged
  html_notebook:
    nature:
      beforeInit: https://platform.twitter.com/widgets.js
---

<!-- progressively evaluate dplyr pipes and ggplot2 viz -->

<!-- <blockquote class="twitter-tweet"><p lang="en" dir="ltr">regular reminder to always check that your code is doing what you think it is<br><br>mine was not cleaning the data the way I thought it was!! ahhH!</p>&mdash; Auriel Fournier (@RallidaeRule) <a href="https://twitter.com/RallidaeRule/status/1182328743202902017?ref_src=twsrc%5Etfw">October 10, 2019</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script> -->


```{r, include=F}
knitr::opts_chunk$set(cache = T, echo = F, warning = F, message = F, cache = T, error = F)
```


## What are flipbooks?  


<!-- # Pedal to the Metal: NSSD 86 1:01 -->
<!-- # "Now that I'm thinking about this, I would so prefer to watch live coding than to go -->
<!-- # to someone's code and try to figure out what's happening in it.   -->
<!-- # I would rather see it built than see the finish product than reverse engineer it." -->

<!-- # NSSD 86: minute 21 -->
<!-- # Let me just go through this as if it is a step-by-step analysis so I can see the  -->
<!-- # intermediate objects ... it almost makes it harder to translate for someone -- ...  -->
<!-- # if I need to do exactly what they were doing then great, but if not I have to rip out  -->
<!-- # the actual code from the function in order to start from scratch. -->

  

<!-- The presentation will guide attendees in creating their own Flipbooks (with Xaringan slides) or mini Flipbooks (gif output).   -->


<!-- "Flipbooks" are a new tool that present side-by-side, aligned, incremental code-output evolution via automated code parsing and reconstruction. -->

<!-- Visual superimposition of new code on previous code and new output on previous output yields a ["blink comparator"](https://en.wikipedia.org/wiki/Blink_comparator){target="_blank"} type experience, and incrementalism assists viewers in linking specific components of code with resultant behavior.  Flipbooks are promising as communication, educational, and reference medium. -->

<!-- Click on the example below. Then you should be able to "flip" (left and right arrows) through the book to see how to make an animated racing bar chart of population size in the most populous countries over time.   -->


<!-- The flipbook is a new tool for dynamically presenting code alongside output.   -->

"Flipbooks" are new tools that present side-by-side, aligned, incremental code-output evolution via automated code parsing and reconstruction.

Good examples of code behavior help programmers accomplish unfamiliar tasks. Tools for communicating code behavior have improved in recent years.  "Dynamic documents" show chunks of code and its resultant output, like a graph or a data table, immediately thereafter. But creators using these tools often must choose between big-picture or narrow-focus demonstration. Creators tend to either demonstrate a complete code pipeline that accomplishes a realistic task or instead demonstrate a minimal example which clarifies the behavior of a particular function, but how it might be used in a larger project isn't clear. Flipbooks help address this problem, allowing creators to build a real work product and allowing consumers to walk incrementally through a code pipeline, focusing on any unfamiliar steps. 

Essentially, *flipbooks* seek to minimize guesswork about how code input and its output relate.  As such flipbooks hold much promise as an educational, communication, and reference medium. 

<!-- The code that assists in creating flipbooks (automating code parsing, reconstruction and presentation) was initially implemented in October 2018.  I've made many improvements and refinements to this parsing code that helps with usability and makes the code more robust, and have written complementary code for the flipbook format.  However, there is still much work to be done to make these tools widely accessible.  I am requesting funds to hire two masters student workers to help in this effort.  Students will help test the flipbook R package.  They will also work on educational materials to demonstrate how to use the code, and will also assist in the maintenance of a flipbook library, describing and showcasing exemplary books - especially those that are useful in clarifying behavior of base R and useful R packages.   -->

<!-- The flipbook building tools parse code in a data manipulation or visualization pipeline and then build it back up incrementally. Aligned superimposition of new code and output atop previous code and output makes it easy to identify how each code addition triggers changes in output.  -->


<!-- Scholars using a reproducible workflow in which they document each step of data manipulation and analysis in code rather than using software with a graphical user interface.   Accomplishing new or unpracticed tasks when using such a programmatic workflow is less burdensome with the availability of good examples.  Tools for communicating examples have improved in recent years.  Especially embraced are tools that show code and its resultant output immediately thereafter --- the case of `Jupytr` notebooks and `Rmarkdown` documents.   Nevertheless, creators  using these tools must choose between big-picture or narrow-focus demonstration; creators tend to either demo a complete code pipeline that accomplishes a realistic task or instead demonstrate a minimal example which makes clear the behavior of a particular function, but how it might be used in a larger project isn't clear. -->

<!-- Flipbooks help address this problem, allowing the creator to present a full demonstration which accomplishes a task, and gives the *user* the opportunity to focus on unfamiliar steps of the work pipeline. Flipbooks parse code in a pipeline of data manipulation or visualization and then build it back up incrementally.  Each of the a sequence of steps can be examined as needed by the viewer. Visual superimposition of new code and output atop previous code and output makes it easy to identify how each code change triggers changes in output. Currently, the implementation is designed to handle the *R* programming language, which is popular in the field of statistical analysis.    -->

<!-- ### In technical terms? -->



<!-- Aligned superimposition of new code on previous code alongside the new output on previous output means that readers identify differences easily via motion detection.  This contrast with comparisons where there is spatial displacement - which imposes a greater cognitive burden. -->

### Example

Click once in the box the below to initiate the session; then use the right and left arrow to "flip" through the book to see how to make an animated "racing" bar chart of the population size of the most populous countries over time.   



<!-- --- -->

<!-- Submission Title -->
<!-- Flipbooks -->
<!-- Learning Objective - What will attendees learn from your talk? -->


<!-- --- -->
<!-- The talk will introduce Flipbooks, a tool for presenting side-by-side, aligned, incremental code-output evolution via automated code parsing and reconstruction. The presentation will also guide attendees in creating their own Flipbooks (Xaringan slides) or mini Flipbooks (gif output). -->
<!-- --- -->

<!-- Submission Abstract -->

<!-- --- -->



```{r}
knitr::include_url("https://evamaerey.github.io/little_flipbooks_library/racing_bars/racing_barcharts.html#1")
```


## Origins

In mid 2018 Garrick Aden-Buie, a data scientist at the Moffitt Cancer Center, created and debuted a side-by-side format for code and figure output in the slide-show tool, Xaringan, an RMarkdown tool.  He demonstrated in incremental steps how someone might build up a plot using the popular R visualization package ggplot2.  In August 2018, I emulated this work, but using a code style I developed which I called "slowggplott", which is entirely sequential and more incremental than typical workflows.  In September, Emi Tanaka, a Lecturer in Statistics at the University of Sydney, implemented a similar sequential workflow demonstration, that was partially automated.  

I approached Garrick and Emi in October about working together to create a "flipbook" of exemplary data visualizations.  We soon had a fully automated framework to parse ggplot2 code, and present its reconstruction along with associated output!  

The procedure to automatically parse and reconstruct code was tested and refined on small projects.  Then, I created a "flipbook" of a sizable collection of my portfolio of data visualizations.  After writing descriptions to introduce the plots, I publicized the work in February, 2019; the flipbook was met an enthusiastic reception.  

Since then, I've taken the development lead in making a number of changes to the foundational code to improve the tool and to extend use cases. 


<!-- - [x] Adding more "connector" operators for pausing code (%>%, ->, %+%) to enable presentation beyond data visualization -->
<!-- - [x] Allowing code wrapping (breaks occur not necessarily at the end of lines) -->
<!-- - [x] Determining partial code completeness based on paretheses balancing -->
<!-- - [x] Using code parser rather than regular expressions to identify role of text segments -->

<!-- https://www.garrickadenbuie.com/blog/decouple-code-and-output-in-xaringan-slides/ -->
<!-- https://emitanaka.rbind.io/post/knitr-knitr-code/ -->

<!-- - the code automating the code-walk through has been updated to "pause" at more "connectors" in the code -->
<!-- - to allow code wrapping -->
<!-- - allow the user to indicate at what point walk-through should be "paused" -->
<!-- - adding balanced parentheses as a condition of   -->

 <!-- The usability of flipbooks is also being improved by creating tools for the creation of visual table of contents, from which users can link to specific visualizations to view code-plot evolution.    -->

<!-- ## Origins - the ideas hose -->

<!-- <blockquote class="twitter-tweet"><p lang="en" dir="ltr">Good ideas! Very <a href="https://twitter.com/hashtag/ggplot?src=hash&amp;ref_src=twsrc%5Etfw">#ggplot</a>. A layered presentation of graphics. <a href="https://t.co/M4tRB9bGS5">https://t.co/M4tRB9bGS5</a></p>&mdash; Gina Reynolds (@EvaMaeRey) <a href="https://twitter.com/EvaMaeRey/status/991269331257487366?ref_src=twsrc%5Etfw">May 1, 2018</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script> -->

<!-- <blockquote class="twitter-tweet"><p lang="en" dir="ltr">I like this implementation even better. I&#39;m using aes() outside of the ggplot() function or a geom function seems not to be conventional. I just tried and see that it works in this case as I was hoping. But, should it be avoided? bad style? <a href="https://twitter.com/StatGarrett?ref_src=twsrc%5Etfw">@StatGarrett</a> <a href="https://t.co/GxUGo1agHU">pic.twitter.com/GxUGo1agHU</a></p>&mdash; Gina Reynolds (@EvaMaeRey) <a href="https://twitter.com/EvaMaeRey/status/992341968125493248?ref_src=twsrc%5Etfw">May 4, 2018</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script> -->


<!-- <blockquote class="twitter-tweet"><p lang="en" dir="ltr">Really dig how <a href="https://twitter.com/grrrck?ref_src=twsrc%5Etfw">@grrrck</a> builds up <a href="https://twitter.com/hashtag/ggplot2?src=hash&amp;ref_src=twsrc%5Etfw">#ggplot2</a> syntax w/ 📊:<br>📽 &quot;A Gentle Guide to the Grammar of Graphics with ggplot2&quot; <a href="https://t.co/2Okhri7Hox">https://t.co/2Okhri7Hox</a> <a href="https://twitter.com/hashtag/rstats?src=hash&amp;ref_src=twsrc%5Etfw">#rstats</a> <a href="https://twitter.com/hashtag/dataviz?src=hash&amp;ref_src=twsrc%5Etfw">#dataviz</a> <a href="https://t.co/qO43IeTtHG">pic.twitter.com/qO43IeTtHG</a></p>&mdash; Mara Averick (@dataandme) <a href="https://twitter.com/dataandme/status/995650384155693061?ref_src=twsrc%5Etfw">May 13, 2018</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script> -->

<!-- https://pkg.garrickadenbuie.com/gentle-ggplot2/#46 -->

<!-- <blockquote class="twitter-tweet"><p lang="en" dir="ltr">Here, building up a <a href="https://twitter.com/hashtag/ggplot2?src=hash&amp;ref_src=twsrc%5Etfw">#ggplot2</a> as slowly as possible, <a href="https://twitter.com/hashtag/rstats?src=hash&amp;ref_src=twsrc%5Etfw">#rstats</a>. Incremental adjustments. <a href="https://twitter.com/hashtag/rstatsteachingideas?src=hash&amp;ref_src=twsrc%5Etfw">#rstatsteachingideas</a> <a href="https://t.co/nUulQl8bPh">pic.twitter.com/nUulQl8bPh</a></p>&mdash; Gina Reynolds (@EvaMaeRey) <a href="https://twitter.com/EvaMaeRey/status/1029104656763572226?ref_src=twsrc%5Etfw">August 13, 2018</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script> -->

<!-- <blockquote class="twitter-tweet"><p lang="en" dir="ltr">Inspired by <a href="https://twitter.com/grrrck?ref_src=twsrc%5Etfw">@grrrck</a> and <a href="https://twitter.com/EvaMaeRey?ref_src=twsrc%5Etfw">@EvaMaeRey</a>, made the kunoichi + ninjutsu (ninja-theme) version of <a href="https://twitter.com/hashtag/ggplot?src=hash&amp;ref_src=twsrc%5Etfw">#ggplot</a> tutorial although Garrick already does explaining this in his excellent blog <a href="https://t.co/msXfg14Ztn">https://t.co/msXfg14Ztn</a>. Gist for ninja-theme here: <a href="https://t.co/soHH4Qvz4F">https://t.co/soHH4Qvz4F</a> <a href="https://twitter.com/hashtag/rstats?src=hash&amp;ref_src=twsrc%5Etfw">#rstats</a> <a href="https://t.co/YlRHAGnaUm">pic.twitter.com/YlRHAGnaUm</a></p>&mdash; Emi Tanaka 🌾 (@statsgen) <a href="https://twitter.com/statsgen/status/1041279648452108289?ref_src=twsrc%5Etfw">September 16, 2018</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script> -->




## Examples

The February 2019 debut of the flipbook technique unveiled in its application to a substantial collection of my data visualization code in [the ggplot flipbook](https://evamaerey.github.io/ggplot_flipbook/ggplot_flipbook_xaringan.html){target="_blank"}.  Shortly thereafter, the reveal code was extended to handle data manipulation pipelines which allowed for the creation of [The Tidyverse in Action](https://evamaerey.github.io/tidyverse_in_action/tidyverse_in_action.html){target="_blank"}.   

Since then flipbook tool has been applied in explaining how to create visualizations important to statistical pedagogy, as in [The Galton Board flipbook](https://evamaerey.github.io/little_flipbooks_library/galton_board/galton_board.html){target="_blank"}
; political decision rules as in [D’Hondt/Jefferson Allocation](https://evamaerey.github.io/little_flipbooks_library/dhondt_jefferson_allocation/dhondt_jefferson_allocation){target="_blank"}; and custom plot types, as in [a racing barchart](https://evamaerey.github.io/little_flipbooks_library/racing_bars/racing_barcharts.html){target="_blank"}.

Returning to the initial focus of data visualization using ggplot, new resources have been created focused on providing systematic and theoretical foundations for understanding ggplot2 -- the implementation of the powerful Grammar of Graphics framework in R -- including [a ggplot grammar guide](https://evamaerey.github.io/ggplot2_grammar_guide/ggplot2_grammar_guide#1){target="_blank"} and [ggplot themes](https://evamaerey.github.io/little_flipbooks_library/taming_themes_in_ggplot/taming_ggplot_themes.html){target="_blank"}.  

Use of the framework has also been explored outside of its ["Tidyverse"](https://www.tidyverse.org/){target="_blank"} roots in [a data.table flipbook](https://evamaerey.github.io/little_flipbooks_library/data.table/data.table){target="_blank"}, and [leaflet's interactive maps for R](https://evamaerey.github.io/little_flipbooks_library/leaflet/leaflet#1){target="_blank"} for example.  

## Testing phase

A code repository for the smaller flipbooks projects "lives" in one place, [here](https://github.com/EvaMaeRey/little_flipbooks_library), for reference.   Larger projects' code is also open access and is available for examination and emulation.

The next frontier was making the technique easy to use, problem free.  One step to this end was to use others' code to create Flipbooks as was done in [#TidyTuesday walk through](https://evamaerey.github.io/tidytuesday_walk_through/tidytuesday_highlights.html){target="_blank"}.  Four visualizers generously allowed me to use their beautiful work.  Problems with the code parsing were uncovered and addressed through this process.

Additionally, I'm worked on and testing materials to help people create their own books with minimal hassle.  A guide exists in the [A flipbook minimal example](https://evamaerey.github.io/little_flipbooks_library/tidytuesday_minimal_example/tidytuesday_minimal_example){target="_blank"} and the [The checkerboard example](https://evamaerey.github.io/little_flipbooks_library/checkerboard/checkerboard.html){target="_blank"}.  In a June 2019 workshop participants at the University of Illinois, successfully produced their own flipbook using the minimal example as a guide.  

In this phase, research assistance from DU masters student Ryan Granier was extremely valuable in helping to move the project forward.  Thanks Ryan!

## Package



#### Modalities

##### Reveal modalities

##### Extension to non-sequential code

Sequentially revealing code is the current implementation of flipbooks. However, there are certainly circumstances in which it would make more sense to reveal things out-of-order to better communicate about code input and subsequent output behavior.  Tackling non-sequential reveal is therefore on the to do-list.   

##### Presentation modalities

UseThere are several modalities for using flipbookr.  These are presented in the following flipbook (which is built from the template delivered with the package):

```{r}
knitr::include_url("https://evamaerey.github.io/little_flipbooks_library/flipbookr/skeleton#1")
```

 in traditional presentations

An additional adaptation of the parsing and reveal code allows users to slowly reveal the components of a plot without also featuring the code to produce the plot in [A flipbook minimal example](https://evamaerey.github.io/little_flipbooks_library/minimal_example/minimal_example){target="_blank"}, a strategy that International Health Professor Hans Rosling used in his stunning and viral presentations, and a technique that is recommended by data journalist Mona Chalabi and Harvard Professor of Government Matthew Blackwell.

The### Strategies for showing evolution


```### Guiding template

## Implementation

```{r}
knitr::include_url("https://evamaerey.github.io/little_flipbooks_library/flipbookr/flipbookr_building_blocks#1")
```

 

<!-- https://medium.com/@urban_institute/4-observations-on-animating-your-data-visualizations-cf987b069c35 Jon Schwabish-->


### Theoretical description

I'm currently working on a manuscript thinking about flipbooks from a theoretical perspective, describing the format and our particular implementation, as well as articulating its merits in reducing the burden in connecting code to its resultant output.  



## Reception

The feedback on the flipbook format has been very encouraging and a big motivation to keep working on this project.  The ggplot flipbook was a [top story on Hacker News](https://news.ycombinator.com/item?id=19356354) and was featured as a highlight in the [R Weekly](https://www.r-craft.org/r-news/r-weekly-2019-07-flipbook-package-building/) newsletter. 


Some of the most encouraging feedback about flipbooks has come from individuals, via Twitter.  Their comments can be viewed [here](https://evamaerey.github.io/little_flipbooks_library/about/cheers_for_flipbooks).

Feedback comes from individuals around the world including the following locations!

```{r, results='asis'}
library(tidyverse)
load("data_products/tweets.Rdata")
df %>% 
  select(location) %>% 
  drop_na() %>% 
  pull(location) %>% 
  cat(sep = " | ")
```



<!-- ```{r, out.width="48%"} -->
<!-- knitr::include_url("https://evamaerey.github.io/ggplot_flipbook/ggplot_flipbook_xaringan.html") -->
<!-- knitr::include_url("https://evamaerey.github.io/tidyverse_in_action/tidyverse_in_action.html") -->
<!-- knitr::include_url("https://evamaerey.github.io/ggplot2_grammar_guide/ggplot2_grammar_guide") -->
<!-- knitr::include_url("https://evamaerey.github.io/tidytuesday_walk_through/tidytuesday_highlights.html") -->
<!-- ``` -->

