---
title: "tuning ggplot themes"
subtitle: "a flipbook | made with Xaringan"
author: "<br><br>Gina Reynolds, May 2019"
date: "<br>"
output:
  xaringan::moon_reader:
    chakra: libs/remark-latest.min.js
    lib_dir: libs
    css: ["kunoichi", "ninjutsu"]
    nature:
      ratio: "16:9"
      highlightLines: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.height = 6, out.width = "100%", comment = " ", cache = T, message = F, warning = F, dpi = 300)

source("https://raw.githubusercontent.com/EvaMaeRey/little_flipbooks_library/master/xaringan_reveal_parenthetical.R")
```


---

# Table of Contents

- [Ugly Theme Choices](#uglyplot)

- [Text Elements](#textelements)
- [Rectangular Elements](#rectangularelements)
- [Line Elements](#lineelements)

- [Legends](#legend)
- [Sundry](#sundry)
- [Facets](#facet)

- [Preset Themes](#presetthemes)

---


# Introduction

This is a focused extension of [the ggplot flipbook](https://evamaerey.github.io/ggplot_flipbook/ggplot_flipbook_xaringan.html).The aim is to illuminate theme statements. 




---

# Base Plot

We'll work with the plot below to apply all the theme adjustments.  In the next slide, we'll build this plot and save it as the plot object `g`.  To this plot object we'll add the theme statements.  You might check out [The Tidyverse in Action](https://evamaerey.github.io/tidyverse_in_action/tidyverse_in_action.html) if you are puzzled about how to build `g`. The [ggplot2 Cheatsheet](https://www.rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf) is also an excellent resource. 

```{r the_plot, eval = T, echo = F}
gapminder %>%
  filter(year == 1992) %>%
  ggplot() +
  aes(x = gdpPercap, y = lifeExp) +
  geom_point(alpha = .8, shape = 21, 
             fill = "white") +
  aes(col = continent, fill = continent) +
  geom_point(alpha = .3, shape = 21) +
  aes(size = pop) +
  scale_size(guide = F) +
  labs(title = "Wealth and expected longevity\n in 1992") +
  labs(subtitle = "Data from gapminder package in R") +
  labs(x = "GDP per Capita") +
  labs(y = "Life Expectency") +
  labs(tag = "Plot 1") +
  labs(col = "Continent", fill = "Continent") +
  labs(caption = "Vis: @EvaMaeRey with ggplot") ->
g
```

```{r, echo = F, out.width="50%"}
g
```



---

`r apply_reveal("the_plot")`





---

# Thematic adjustment overview

--

## Components

The components in a ggplot graph are usually things you can point to on a plot.  

For example the *legend box*, or the *y axis title*, or the *plot title*.   


--

## Elements

These components have a variety of elements.  For example the legend box has the perimeter of the box (the line) and it has the area of the box associated with it.  To "point" to specify the element parts of the components, functions like element_line(), element_rect(), and element_text() are used.  To delete an entire component, element_blank() can be used.

--

## Aesthetics

Aesthetics assignment for theme elements are similar to those available for geometric shapes in the plot. Aesthetics include color, size, line_type, and so on. 


---

# Graph Components

So, what components can be adjusted? (there are lots!)

> aspect.ratio

> axis.title, axis.title.x,
  axis.title.x.top, axis.title.x.bottom, axis.title.y, axis.title.y.left,
  axis.title.y.right, axis.text, axis.text.x, axis.text.x.top,
  axis.text.x.bottom, axis.text.y, axis.text.y.left, axis.text.y.right,
  axis.ticks, axis.ticks.x, axis.ticks.x.top, axis.ticks.x.bottom,
  axis.ticks.y, axis.ticks.y.left, axis.ticks.y.right, axis.ticks.length,
  axis.line, axis.line.x, axis.line.x.top, axis.line.x.bottom, axis.line.y,
  axis.line.y.left, axis.line.y.right
  
> legend.background, legend.margin,
  legend.spacing, legend.spacing.x, legend.spacing.y, legend.key,
  legend.key.size, legend.key.height, legend.key.width, legend.text,
  legend.text.align, legend.title, legend.title.align, legend.position,
  legend.direction, legend.justification, legend.box, legend.box.just,
  legend.box.margin, legend.box.background, legend.box.spacing
  
> panel.background, panel.border, panel.spacing, panel.spacing.x,
  panel.spacing.y, panel.grid, panel.grid.major, panel.grid.minor,
  panel.grid.major.x, panel.grid.major.y, panel.grid.minor.x,
  panel.grid.minor.y, panel.ontop 
  
> plot.background, plot.title,
  plot.subtitle, plot.caption, plot.tag, plot.tag.position, plot.margin
  
> strip.background, strip.background.x, strip.background.y,
  strip.placement, strip.text, strip.text.x, strip.text.y,
  strip.switch.pad.grid, strip.switch.pad.wrap



---
name: uglyplot

# An ugly plot, for locating components!

Now we'll create this plot to get a sense of how thematic modifications can be made, by 1) specifying the component to be adjusted 
2) the element of the component to be adjusted and 3) indicating the aesthetic adjustment.  

The adjustments made should make clear what components of the plot are called in ggplot. 

Later we'll walk through a more focused adjustments.  We'll go through time elements one at at time: text, rectangular, line, etc.  

```{r custom_theme, echo = F, eval = T, out.width="50%"}
g + 
  theme(rect = element_rect(fill = "snow3")) +
  theme(text = element_text(family = "Times New Roman")) +
  theme(axis.title = element_text(color = "slateblue4")) +
  theme(title = element_text(color = "seagreen")) +
  theme(axis.line = element_line(color = "orange")) +
  theme(legend.key = element_rect(fill = "violet")) +
  theme(axis.ticks = element_line(color = "blue")) +
  theme(axis.ticks.x = element_blank()) +
  theme(panel.background = element_rect(fill = "steelblue")) +
  theme(panel.grid = 
          element_line(size = 3, color = "thistle1")) +
  theme(legend.background = element_rect(fill = "goldenrod1")) +
  theme(legend.text = element_text(size = 7)) +
  theme(legend.position = "bottom") +
  theme(legend.margin = margin(t = 0, r = 20, 
                               b = 0, l = 40, unit = "pt")) +
  theme(aspect.ratio = .5) 
```

---


`r apply_reveal("custom_theme")`





---

# What are the elements of these components?

Now we'll focus on different elements of the components: 

- element_line()
- element_rect()
- element_text()

Additionally, element_blank(), allows you to indicate that you don't want the component included in the plot at all. 


---


# Across-the-board adjustment

Across-the-board adjustment to all components with given elements can also be done using the *line*, *rect*, *text*, and *title* adjustments.
These will apply generally where no more specific adjustment has been made.  

> line

> rect

> text 

> title


---



```{r, message = F, warning=F, eval = T}
options(scipen = 10) # adjust when scientific notation turns on
library(tidyverse)
library(gapminder)
```









---

name: textelements

# Focus on text elements!



```{r text_components, eval = T, echo = F, out.width="60%"}
g + 
  theme(plot.tag = 
          element_text(color = "plum4", size = 25)) +
  theme(plot.title = 
          element_text(color = "plum4", 
                       size = 20,
                       lineheight = 2.5)) +
  theme(plot.subtitle = element_text(color = "plum4")) +
  theme(axis.title.y = 
          element_text(color = "plum4", 
                       size = rel(3))) +
  theme(axis.text.y = element_text(color = "plum4")) +
  theme(legend.title = 
          element_text(color = "plum4", size = 22)) +
  theme(legend.text = element_text(color = "plum4")) +
  theme(axis.title.x = element_text(color = "plum4")) +
  theme(axis.text.x = element_text(color = "plum4")) +
  theme(plot.caption = 
          element_text(color = "plum4", size = 20)) +
  theme(text = element_text(family = "Times")) + 
  theme(text = element_text(face = "bold.italic")) +
  theme(text = element_text(
    margin = margin(t = 3, r = 3, b = 3, l = 3, 
                    unit = "pt"))) +
  theme(text = element_text(angle = 2)) +
  theme(plot.title = element_text(hjust = .5)) +
  theme(axis.title.x = element_text(hjust = 1)) 
```

---

`r apply_reveal("text_components")`


---

name: rectangularelements

# Focus on rectangular elements!


```{r rect_components, eval = T, echo = F, out.width="60%"}
g +
  theme(plot.background = 
          element_rect(fill = "purple", 
                       color = "darkolivegreen", 
                       size = 3)) +
  theme(panel.background = 
          element_rect(fill = "purple", 
                       color = "darkolivegreen",
                       size = 3)) +
  theme(legend.background = 
          element_rect(fill = "purple", 
                       color = "darkolivegreen", 
                       size = 3)) +
  theme(legend.key = 
          element_rect(fill = "purple", 
                       color = "darkolivegreen", 
                       size = 3)) +
  theme(panel.border = 
          element_rect(color = "magenta", 
                       fill = NA, size = 4))
```

---

`r apply_reveal("rect_components")`

---

name: lineelements

# Focus on line elements!


```{r line_components, eval = T, echo = F, out.width="60%"}
g +
  theme(axis.ticks.y = 
          element_line(size = 10, 
                       color = "blue")) +
  theme(axis.ticks.x = 
          element_line(size = 10, 
                       color = "blue")) +
  theme(panel.grid.major.y = 
          element_line(size = 2,
                       linetype = "dashed",
                       color = "blue",
                       lineend = "round")) +
  theme(panel.grid.major.x = 
          element_line(size = 2, 
                       arrow = arrow(),
                       color = "blue",
                       lineend = "round")) +
  theme(panel.grid.minor.y = 
          element_line(size = 1, 
                       linetype = "dotted",
                       color = "blue",
                       lineend = "butt")) +
  theme(panel.grid = 
          element_line(color = "blue")) +
  theme(axis.line.x.top = 
          element_line(size = 2,
                       color = "blue")) +
  theme(panel.ontop = TRUE,   
        panel.background = element_blank()) # in front of the data
```

---

`r apply_reveal("line_components")`



---

# What aesthetics of these elements can be adjusted?

| aesthetic | adjustable for element of type: |
|-----|----|
| fill | *rect*  |
| color | *rect, line, text* |
| size |  *rect, line, text* |
| linetype | *rect, line* |
| lineend | *line* |
| arrow | *line* |
| family (font) | *text* |
| face | *text* |
| hjust | *text* |
| vjust | *text* |
| angle  | *text* |
| lineheight |  *text* |
| margin | *text* |
| debug | *text* |
| inherit.blank | *rect, line, text* |



---

name: legends

# More legend options!




```{r legend, eval = T, echo = F, out.width="60%"}
g +
  theme(legend.position = "bottom") +
  theme(legend.position = "none") +
  theme(legend.position = c(.8, .25)) +
  theme(legend.direction = "horizontal") +
  theme(legend.key = element_rect()) +
  guides(fill = guide_legend(size = 35))
```
---

`r apply_reveal("legend")`

---

name: sundries

# More modifications! (Sundries)


```{r sundries, eval = T, echo = F, out.width="60%"}
g +
  theme(plot.tag.position = "topright") + 
  scale_color_discrete(guide = FALSE) +
  scale_fill_discrete(guide = FALSE) +
  scale_x_continuous(position = "top") +
  scale_y_continuous(position = "right") +
  theme(aspect.ratio = .5) 
```

---

`r apply_reveal("sundries")`


---

name: facet

# Focus on facet elements!


```{r small_multiples, eval = T, echo = F, out.width="60%"}
g + 
  facet_wrap(~ continent, 
             strip.position = "right") +
  theme(strip.background = 
          element_rect(fill = "pink")) +
  theme(strip.text.y = 
          element_text(color = "snow", size = 12)) +
  theme(strip.text.y = 
          element_text(angle = -80)) +
  theme(panel.spacing.x = 
          unit(1.5, "lines")) +
  theme(panel.spacing.y = 
          unit(2, "lines")) +
  theme(strip.switch.pad.grid = 
          unit(0, "cm"))
```

---

`r apply_reveal("small_multiples")`

---

name: presetthemes

# "base" ggplot preset themes

Next we'll have at different preset themes available in ggplot that can be applied in one line of code. 

---

```{r theme_presets, eval = F, echo = F}
g + 
  theme_bw() +
  theme_classic() +
  theme_dark() +
  theme_gray() + 
  theme_light() +
  theme_linedraw() +
  theme_minimal() +
  theme_void()
```

---

`r apply_reveal("theme_presets")`

---

# Outside themes package 

Other packages are designed to complement ggplot in providing more these

Featured here are `ggthemes`, `ggpomological`, and `ggdark`.

```{r}
# If you don't have devtools installed
# install.packages("devtools")
# devtools::install_github("gadenbuie/ggpomological")
```

```{r }
library(ggthemes)
library(ggpomological)
library(ggdark)
```

Let's check some of these out!

---

# ggthemes

---

```{r ggthemes, eval = F, echo = F}
g + 
  ggthemes::theme_base() +
  ggthemes::theme_calc() + 
  ggthemes::theme_economist() +
  ggthemes::theme_economist_white() +
  ggthemes::theme_excel() +
  ggthemes::theme_excel_new() +
  ggthemes::theme_few() +
  ggthemes::theme_fivethirtyeight() +
  ggthemes::theme_gdocs() +
  ggthemes::theme_hc() +
  ggthemes::theme_igray() +
  ggthemes::theme_map() +
  ggthemes::theme_solarized() +
  ggthemes::theme_solarized_2() + 
  ggthemes::theme_solid() +
  ggthemes::theme_stata() +
  ggthemes::theme_wsj() 
```

---


`r apply_reveal("ggthemes")`

---



---

# ggpomological

---

```{r ggpomological, echo = F, eval = F}
g +
  ggpomological::theme_pomological() +
  ggpomological::theme_pomological_nobg() +
  ggpomological::theme_pomological_plain() 
```

`r apply_reveal("ggpomological")`

---

# ggdark


---

```{r ggdark, echo = F, eval = F}
g +
  ggdark::dark_theme_bw() +
  ggdark::dark_theme_classic() +
  ggdark::dark_theme_dark() +
  ggdark::dark_theme_gray() +
  ggdark::dark_theme_light() +
  ggdark::dark_theme_linedraw() +
  ggdark::dark_theme_minimal() +
  ggdark::dark_theme_test()
```

`r apply_reveal("ggdark")`




-------







```{css, eval = TRUE, echo = F}
.remark-code{line-height: 1.5; font-size: 70%}
```




<!-- # hrbrthemes -->

<!-- --- -->

<!-- ```{r} -->
<!-- g +  -->
<!--   hrbrthemes::theme_ft_rc() + -->
<!--   hrbrthemes::theme_ipsum() + -->
<!--   hrbrthemes::theme_ipsum_ps() + -->
<!--   hrbrthemes::theme_ipsum_pub() + -->
<!--   hrbrthemes::theme_ipsum_rc() + -->
<!--   hrbrthemes::theme_ipsum_tw() + -->
<!--   hrbrthemes::theme_modern_rc() -->
<!-- ``` -->




