---
title: "Minimal Galton Board"
author: "Gina Reynolds"
date: "6/29/2019"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["kunoichi", "ninjutsu"]
    nature:
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r, message = F, warning = F}
source("https://raw.githubusercontent.com/EvaMaeRey/little_flipbooks_library/master/xaringan_reveal_parentheses_balanced.R")
```


```{r setup, include=F}
knitr::opts_chunk$set(fig.width = 6, message = F, warning = F, 
                      comment = "", cache = F)
```




```{r}
library(tidyverse)
options(gganimate.nframes = 400, scipen = 10)
```

---

# Define paramters

```{r}
n <- 70  # number of ball bearnings 
stop_level <- 10 # number of perturbation levels 
                # make it an even number
levels <- 44 # greater than stop_levels
```


---

# Write some code to describe the random-walk of ball bearings

---

```{r ball_bearings, eval = T, echo = F}
set.seed(2019)
crossing(unit_id = 1:n, 
         level = 1:levels - 1) %>% 
  mutate(perturbation =     # moves
           sample(c(-1,1),  # left or right
                  n(), # each ball at each level 
                  replace = T)) %>%
  group_by(unit_id) %>% # operations on each ball
  mutate(perturbation = 
           ifelse(row_number() == 1, 
                  yes = 0, # start centered
                  no = perturbation)) %>% 
  # each ball should release one at a time
  mutate(time = # displacing them in time w/ 
           row_number() + 
           # using unit id
           unit_id * 3 - 1) %>% 
  filter(time > 0) %>% 
  mutate(x_position = # we get the x position
           # by summing the cumulative distributions
           cumsum(perturbation)) %>% 
  # if ball is beyond the perturbation levels
  mutate(x_position = # we overwrite the x position
           ifelse(level <= stop_level,
                  yes = x_position, 
                  no = NA)) %>% 
  # then fill in with the last x position
  fill(x_position) %>% 
  ungroup() ->
ball_bearings
```



---

`r apply_reveal("ball_bearings")`

---

# Now some code to count the number of ball bearings seen at the final level.

We want to keep track of how many ball bearings end up in each position.

---

```{r collect_ball_bearings, eval = T, echo = F}
ball_bearings %>% 
  filter(level == (levels - 1) ) %>% 
  rename(final_time = time) %>% 
  crossing(time = as.numeric(1:max(ball_bearings$time))) %>% 
  group_by(time, x_position) %>% 
  summarise(x_position_count = sum(time > final_time)) ->
ball_bearings_collect
```

`r apply_reveal("collect_ball_bearings")`


---

```{r architecture, eval = T, echo = F}
# Lets make some pegs
crossing(unit_id = 
           -stop_level:stop_level, 
         level = 1:stop_level) %>% 
  mutate(transparent = 
           (unit_id + level) %% 2) ->
pegs
# Lets make walls
crossing(unit_id = 
           -(stop_level + 1):(stop_level + 1), 
         level = stop_level:levels) %>% 
  mutate(transparent = 
           unit_id %% 2) ->
walls
```


`r apply_reveal("architecture")`


---

```{r}
ball_bearings_size <- 2
peg_size <- 3
```

---

```{r, static_build, eval = T, echo = F}
ggplot(ball_bearings) +
  aes(y = level) +
  aes(x = x_position) +
  scale_y_reverse() +
  aes(group = unit_id) +
  geom_point(data = walls, 
             aes(x = unit_id, alpha = transparent), 
             col = "grey30", size = peg_size) +
  geom_point(data = pegs, 
             aes(x = unit_id, alpha = transparent), 
             col = "grey30", size = peg_size) +
  geom_segment(x = -sqrt(n), xend = -1.5, 
               y = 0, yend = 0) +
  geom_segment(x = sqrt(n), xend = 1.5, 
               y = 0, yend = 0) +
  geom_abline(intercept = 1.5, 
              slope = -1) +
  geom_abline(intercept = 1.5, 
              slope = 1) +
  annotate(geom = "tile", 
           height = 2, width = 2, 
           x = 0 , y = -1.5) +
  annotate(geom = "tile", 
           height = 2, width = 1.75, 
           x = 0 , y = -1.5, fill = "white") +
  geom_rect(data = ball_bearings_collect,
            mapping = aes(xmin = x_position - .35, 
                          xmax = x_position + .35,
                          ymax = max(ball_bearings$level) + 1 - x_position_count*1.5,
                          ymin = max(ball_bearings$level) + 1,
                          group = x_position, 
                          y = NULL, x = NULL),
            fill = "darkgrey") +
  geom_point(size = ball_bearings_size, 
             col = "steelblue") +
  coord_equal() +
  geom_hline(yintercept = stop_level, 
             linetype = "dotted") +
  scale_alpha_continuous(range = c(0, 1), guide = F) +
  theme_void() ->
g
```

---

`r apply_reveal("static_build")`

---

```{r, animation, eval = F, echo = F}
g +  
  gganimate::transition_time(time = time) + 
  gganimate::shadow_wake(wake_length = .05) + 
  gganimate::ease_aes("bounce-in-out") #REVEAL
```

`r paste(knitr::knit(text = partial_knit_chunks("animation", show_code = T, user_reveal_defined = T)), collapse = "\n")`

---

# Binomial distribution

Now the fast way to simulating...  

---

```{r distribution, eval = F, echo = F}
set.seed(1999)
n_balls <- 1000
n_perturb <- 200
crossing(ball_id = 1:n_balls, 
         1:n_perturb) %>% 
  mutate(perturbation = 
           sample(c(-1,1), 
                  n(), 
                  replace = T)) %>% 
  group_by(ball_id) %>% 
  summarise(x_position = 
              sum(perturbation)) %>% 
  ggplot() +
  aes(x = x_position) +
  geom_histogram(binwidth = 2, 
                 col = "lightgrey",
                 fill = "steelblue",
                 alpha = .7) +
  theme_minimal()
```

`r apply_reveal("distribution")`










```{css, eval = TRUE, echo = F}
.remark-code{line-height: 1.5; font-size: 50%}
```
