---
title: "A Minimal Flipbook"
author: "Gina Reynolds, May 2019"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["kunoichi", "ninjutsu"]
    nature:
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r setup, include=F}
knitr::opts_chunk$set(fig.width = 6, message = F, warning = F, 
                      comment = "", cache = F)
options(tibble.print_max = Inf) 
```




---

# Introduction

This is a minimal example to demonstrate how to create a flipbook that will walk through data wrangling and plots pipelines made with the Tidyverse.  The functions that make this possible are the work of Emi Tanaka, Garrick Aden-Buie and myself, and are built for Xaringan, an Rmarkdown file type for creating presentation slides; the functions make use of the function `knitr:::knit_code$get()`.

Interested in more flipbooks? Check out 

- [the ggplot flipbook](https://evamaerey.github.io/ggplot_flipbook/ggplot_flipbook_xaringan.html)
- [The Tidyverse in Action](https://evamaerey.github.io/tidyverse_in_action/tidyverse_in_action.html)

For more about Xaringan:

- [Xaringan presentation slides](https://slides.yihui.name/xaringan/)

The sequential workflow of the Tidyverse makes incremental display of pipelines ideal:  

- [www.tidyverse.org](https://www.tidyverse.org/)


---

# Set up

Okay. Let's load the the `reveal for xaringan` functions for "flipbooking" and the `tidyverse`.

```{r}
if(file.exists("../xaringan_reveal_parenthetical.R")){

  source(file = "../xaringan_reveal_parenthetical.R")
  
} else {

source(file = "https://raw.githubusercontent.com/EvaMaeRey/little_flipbooks_library/master/xaringan_reveal_parenthetical.R")
  
}


library(tidyverse)
library(gapminder)
```

```{r}
gapminder_2007 <- 
  gapminder %>%
  filter(year == 2007)
```




---

# Where we are going: 

We'll create this plot.  I have **echo** set to FALSE in the code chunk options here so that you don't see the code, and **eval** to TRUE so that the plot output is produced.  The code chunk is given the name "gapminder2007", and this is used in the in apply_reveal function, which breaks down code by wrangling and plot statements. 

```{r gapminder2007, echo = F, eval = T, fig.height=5}
ggplot(data = gapminder_2007) + 
  aes(x = gdpPercap) +  # x axis 
  aes(y = lifeExp) +  # y axis
  geom_point() +
  aes(color = continent)
```
 

 
---

# How do we get there?

In the next slide, we'll walk through the code that produces this plot, and the output along the way. We use the code `apply_reveal("gapminder2007")` in-line to access the code from the code chunk called *gapminder2007*.
 
---

`r apply_reveal("gapminder2007")`


---

# Other pipeline connectors

The reveal functions are also written to recognize the pipe, `%>%`, and reverse assignment operator `->`, at the end of a line as reveal break points.  An example follows. 

```{r gapminder2007pipe, echo = F, eval = T, fig.height=5}
gapminder %>% 
  filter(year == 2007) %>% 
  ggplot() + 
  aes(x = gdpPercap) + 
  aes(y = lifeExp) + 
  geom_point() +
  aes(col = continent) -> 
  g
```


---

`r apply_reveal("gapminder2007pipe")`

---

# Hide code

You might be interested in showing the evolution of a plot, without focusing on the code. Let's see an example of this.  

Consider the case where you want to show the x-axis, then add the y-axis, show one data point, and then add the rest of the data.  

Building up plots slowly in presentations has been a technique employed by Hans Rosling and Mona Chalabi.  Matthew Blackwell tweeted this advice in 2018 -- it resonated with a lot of folks!

---

First we prep a lot of the plot.  In a presentation we would set echo to FALSE. 

```{r}
gapminder_2007 <- gapminder %>% 
  filter(year == 2007) 
g <- ggplot(gapminder_2007) + 
  theme_minimal(base_size = 22)
```

---

Then we prep a code chunk so that lines end where we want breaks.  


```{r, gapminder2007plotonly, eval = F}
g + aes(x = gdpPercap) + labs(x = "GDP per Capita ($US)") +
  aes(y = lifeExp) + geom_blank() + labs(y = "Life Expectency") +
  geom_point(data = gapminder_2007 %>% filter(country == "Spain")) +
  geom_point() 
```

Then we'll use the name of this code chunk in a reveal function "partial_knit_chunks".  

---

`r paste(knitr::knit(text = partial_knit_chunks("gapminder2007plotonly", show_code = F)), collapse = "\n")`



```{css, eval = TRUE, echo = F}
.remark-code{line-height: 1.5; font-size: 100%}
```


