---
title: "diff-in-diff vis w ggplot"
author: "Evangeline Reynolds"
date: "5/18/2018"
output:
  xaringan::moon_reader:
    chakra: libs/remark-latest.min.js
    lib_dir: libs
    css: ["kunoichi", "ninjutsu"]
    nature:
      ratio: "16:9"
      highlightLines: true
---

Using a difference-in-difference framework to look at the effect of policy interventions is a popular research design.  A binary exlanatory variable that turns on and stays on lends itself to ease of interpretation!

In the context of such a design, what is the essential visual inspections?  Below I suggest four plots that researchers and their audiences may find useful in visually inspecting the timing of interventions and the relationships between a policy intervention and outcomes.

I use ggplot2 to implement the visualization. The ability to overwrite global aesthetics, using the aes() function, means that we move from one plot to another with little additional code.  A little over a year ago I learned about declaring the aes() on it's own line and maybe novelty bias is at work here, but I find the capability to be a lot of fun to play with!


---

The five plots are as follows:

- two plots showing the timing of intervention and the cross-sectional cases (one aligning the interventions)
- two plots showing how the intervention relates to the response variable of interest (one aligning the interventions)
- a plot that breaks up the data into individual time-series for the response variable for each cross-sectional unit

Another option for visualizing such data is using the new package [panelView](http://yiqingxu.org/software/panelView/panelView.html), which certainly gave me additional inspiration for this exercise and looks useful!  


```{r}
library(tidyverse)
library(gapminder)
source("../xaringan_reveal_parenthetical.R")
```




```{r}
min_year <- min(gapminder$year)
max_year <- max(gapminder$year)
span <- max_year- min_year

gapminder %>% 
  select(country) %>% 
  distinct() %>% 
  sample_frac(.5) %>% 
  mutate(intervention_year = 
           runif(n = n(), min = min_year - span, 
                 max = max_year)) %>% 
  mutate(intervention_year = round(intervention_year)) ->
synthetic_interventions
```

```{r}
synthetic_interventions %>% 
  full_join(gapminder) %>% 
  mutate(treatment = case_when(year >= intervention_year ~ "treated",
                             year < intervention_year ~ "not treated",
                             is.na(intervention_year) ~ "not treated")) %>%
  group_by(country) %>% 
  mutate(mean_treated = mean(treatment == "treated")) %>% 
  arrange(mean_treated) %>% 
  ungroup() %>% 
  mutate(country = fct_inorder(country)) -> 
panel_prepped
```


```{r essential_plots}
ggplot(panel_prepped) +
  aes(x = year, y = country, group = country, col = treatment) +
  labs(caption = "Four essential plots for diff-in-diff | Gina Reynolds") +
  geom_line(alpha = .5) +
  geom_point(size = 1.5) +
  theme_minimal() +
  scale_color_manual(values = c("lightgrey", "steelblue4" )) +
  labs(y = "country") +
  labs(title = "Plot 1: Treatment-control summary") +
  labs(title = "") +
  aes(y = gdpPercap) + 
  scale_y_log10() +
  labs(y = "GDP per cap") +
  labs(title = "Plot 2: Treatment-control response overlaid summary") +
  labs(title = "") +
  aes(x = year - intervention_year) +
  labs(title = "Plot 3: Treatment-control response overlaid summary, aligned intervention") +
  labs(title = "")  +
  aes(x = year) + 
  scale_y_continuous() +
  labs(title = "Plot 4: Treatment-Response faceted")# %+%
  # (panel_prepped %>%
  #    inner_join(panel_prepped %>% select(country) %>% distinct %>% sample_n(5))) +
  # facet_wrap(~ as.character(country), scales = "free_y")
```


`r apply_reveal("essential_plots")`


```{css, eval = TRUE, echo = F}
.remark-code{line-height: 1.5; font-size: 50%}
```

