---
title: "A leaflet flipbook"
author: "Gina Reynolds, June 2019"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["kunoichi", "ninjutsu"]
    nature:
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r}
# Load leaflet R package
library(leaflet)
# This flipbook was created w/ xaringan reveal functions:
source(file = "https://raw.githubusercontent.com/EvaMaeRey/little_flipbooks_library/master/xaringan_reveal_parenthetical.R")

```


---

# Discover Auckland (Birthplace of R)


```{r auckland, eval = T, echo = F}
leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lng = 174.768, 
             lat = -36.852, 
             popup = "The birthplace of R") %>% 
  setView(lng = 174.768, 
          lat = -36.852,
          zoom = 4)
```

---

`r apply_reveal("auckland")`


---

# Discover Denver 

- Explore the provider tiles - full opacity 
- Open maps seem not quite keeping up with DU construction

```{r mymap, eval = T, echo = F}
leaflet() %>% # Leaflet map widget
  addTiles() %>% # OpenStreetMap default
  setView(lat = 39.6759427, 
          lng = -104.9621703, 
          zoom = 15) %>% 
  addProviderTiles(provider =
                     providers$OpenStreetMap.BlackAndWhite) %>% 
  addMarkers(lat = 39.6759427,
             lng = -104.9621703, 
             popup = "Korbel School of International Studies") %>% 
  addProviderTiles(provider =
                     providers$Stamen.Toner) %>% 
  addProviderTiles(provider =
                     providers$OpenTopoMap) %>%
  addCircles(lat = 39.6759427,
             lng = -104.9621703,
             radius = 500,
             color = "orange",
             fillColor = "red") %>% 
  setView(lat = 39.6759427,
          lng = -104.9621703,
          zoom = 8) %>% 
  setView(lat = 39.6759427,
          lng = -104.9621703,
          zoom = 4)
```


---

`r apply_reveal("mymap")`

---



# Discover Ames, Iowa


- a weather layer
- a semi opaque layer



```{r weather, eval = T, echo = F}
leaflet() %>% 
  addTiles() %>% 
  setView(-93.3712, 42.0205, zoom = 7.5) %>%
  addWMSTiles(baseUrl = "http://mesonet.agron.iastate.edu/cgi-bin/wms/nexrad/n0r.cgi",
    layers = "nexrad-n0r-900913",
    options = WMSTileOptions(format = "image/png", 
                             transparent = TRUE),
    attribution = "Weather data © 2012 IEM Nexrad") %>% 
  setView(-93.6, 42.0205, zoom = 11) %>%
  addProviderTiles(providers$OpenRailwayMap,
    options = providerTileOptions(opacity = 0.75))
```


---

`r apply_reveal("weather")`


---

# Discover Boston

Walk-through of almost verbatim example from https://rstudio.github.io/leaflet/basemaps.html

- provider lines and labels

```{r boston, eval = T, echo = F}
leaflet() %>% 
  setView(lng = -71.0589, lat = 42.3601, 
          zoom = 12) %>% 
  addProviderTiles(providers$MtbMap) %>%
  addProviderTiles(providers$Stamen.TonerLines,
    options = providerTileOptions(opacity = 0.35)) %>%
  addProviderTiles(providers$Stamen.TonerLabels)
```


---

`r apply_reveal("boston")`


```{css, eval = TRUE, echo = F}
.remark-code{line-height: 1.5; font-size: 65%}
```

