---
title: "A leaflet flipbook"
author: "Gina Reynolds, June 2019"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["kunoichi", "ninjutsu"]
    nature:
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---



# Introduction

Leaflet is an open source JavaScript library for embedding custom interactive maps.  Leaflet for R "makes it easy to create Leaflet maps from R."  There is a really helpful [website]( https://rstudio.github.io/leaflet/) introducing this package.  And this [flipbook](https://github.com/EvaMaeRey/little_flipbooks_library) draws a lot from its examples, and is intended to further clarify in-between steps.  

Leaflet is definitely a fun package to explore!   


---

```{r}
# Load leaflet R package
library(leaflet)
# This flipbook was created w/ xaringan reveal functions:
source(file = "https://raw.githubusercontent.com/EvaMaeRey/little_flipbooks_library/master/xaringan_reveal_parenthetical.R")

```




---

# Discover Auckland (Birthplace of R)


```{r auckland, eval = T, echo = F}
leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lng = 174.768, 
             lat = -36.852, 
             popup = "The birthplace of R") %>% 
  setView(lng = 174.768, 
          lat = -36.852,
          zoom = 4)
```

---

`r apply_reveal("auckland")`


---

# Discover Denver 

- Explore the provider tiles - full opacity 
- Open maps seem not quite keeping up with DU construction

```{r mymap, eval = T, echo = F}
leaflet() %>% # Leaflet map widget
  addTiles() %>% # OpenStreetMap default
  setView(lat = 39.6774101, 
          lng = -104.9621703, 
          zoom = 15) %>% 
  addProviderTiles(provider =
                     providers$OpenStreetMap.BlackAndWhite) %>% 
  addMarkers(lat = 39.6774101,
             lng = -104.9621703, 
             popup = "University of Denver") %>% 
  addProviderTiles(provider =
                     providers$Stamen.Toner) %>% 
  addProviderTiles(provider =
                     providers$OpenTopoMap) %>%
  addCircles(lat = 39.6774101,
             lng = -104.9621703,
             radius = 500,
             color = "orange",
             fillColor = "red") %>% 
  setView(lat = 39.6774101,
          lng = -104.9621703,
          zoom = 8) %>% 
  setView(lat = 39.6774101,
          lng = -104.9621703,
          zoom = 4)
```


---

`r apply_reveal("mymap")`

---



# Discover Ames


- a semi opaque layer
- adding a rectangle

---




```{r ames, eval = T, echo = F}
leaflet() %>% 
  addTiles() %>% 
  setView(-93.625, 42.0205, zoom = 11) %>%
  addRectangles(
    lng1 = -93.625 - .03, lat1 = 42.0205 - .03,
    lng2 = -93.625 + .03, lat2 = 42.0205 + .03,
    fillColor = "transparent"
  ) %>% 
  addProviderTiles(providers$OpenRailwayMap,
    options = providerTileOptions(opacity = 0.75)) 
```


---

`r apply_reveal("ames")`


---

# Discover Boston

Walk-through of almost verbatim example from https://rstudio.github.io/leaflet/basemaps.html

- provider lines and labels
- multiple markers
- weather layer
- including external data

---


```{r}
new_england <- read.csv(textConnection("City,Lat,Long,Pop
Boston,42.3601,-71.0589,645966
Hartford,41.7627,-72.6743,125017
New York City,40.7127,-74.0059,8406000
Philadelphia,39.9500,-75.1667,1553000
Pittsburgh,40.4397,-79.9764,305841
Providence,41.8236,-71.4222,177994"))
```


```{r boston, eval = T, echo = F}
new_england %>% 
leaflet() %>% 
  addProviderTiles(providers$MtbMap) %>%
  setView(lng = -71.0589, lat = 42.3601, 
          zoom = 12) %>% 
  addProviderTiles(providers$Stamen.TonerLines,
    options = providerTileOptions(opacity = 0.35)) %>%
  addProviderTiles(providers$Stamen.TonerLabels) %>%
  setView(-84, 42.0205, zoom = 4) %>% 
  addWMSTiles(baseUrl = "http://mesonet.agron.iastate.edu/cgi-bin/wms/nexrad/n0r.cgi",
    layers = "nexrad-n0r-900913",
    options = WMSTileOptions(format = "image/png", 
                             transparent = TRUE),
    attribution = "Weather data © 2012 IEM Nexrad") %>%
  addCircles(lng = ~ Long, lat = ~ Lat, 
             weight = 1, 
             radius = ~ sqrt(Pop) * 30, 
             popup = ~ City, color = "red")  %>%
  addMarkers(lng = ~ Long, lat = ~ Lat) 
```


---

`r apply_reveal("boston")`


---


# Discover Germany

- adding polygons

Somehow the "flipbook" function `apply_reveal()` failed here.  Nevertheless, I'm including the code and the complete result. 

```{r, eval = F}
download.file("https://www.bundeswahlleiter.de/dam/jcr/f92e42fa-44f1-47e5-b775-924926b34268/btw17_geometrie_wahlkreise_geo_shp.zip", "btw17_geometrie_wahlkreise_geo_shp.zip")
unzip("btw17_geometrie_wahlkreise_geo_shp.zip")
```    

```{r, warning=F, message=F}
library(rgdal)
german_districts <- readOGR(dsn = "Geometrie_Wahlkreise_19DBT_geo.shp", layer = "Geometrie_Wahlkreise_19DBT_geo")
```
---

```{r bundestag, eval = T, echo = T}
german_districts %>% leaflet() %>% 
  addTiles() %>% 
  setView(lng = 7, lat = 50, zoom = 5) %>% 
  addPolygons(color = "grey", weight = 1, 
              smoothFactor = 0.5,
              fillOpacity = 0.25,
              fillColor = "blue",
              highlightOptions = 
                highlightOptions(color = "white", 
                                 weight = 2,
                                 bringToFront = TRUE)) ->
g
```

---

```{r}
g
```



---

```{css, eval = TRUE, echo = F}
.remark-code{line-height: 1.5; font-size: 65%}
```

